---
title: "Data Exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 11, warning = FALSE)
```

What kind of information holds the training set? For what does each column stend for? Are there any false observations, anomalies? 

From the names of the columns and the first few rows, we get, that the training set consist of historical purchases of different product from different customers:

* the id of the order (order_id)
* the id of the customer (contact_id)
* the id of the product (product_id)
* date of purchase (purchase_date)
* quantity of product in order (quantity)
* the amount paid (sales_amount)


```{r message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(lubridate)

source("R/utils.R")

purchases <- loadPurchases()
summary(purchases)
head(purchases, 20)
```

Observing the summary, we can see, that the purchases roughly occur in a consistent basis and usually, the customers purchase 1 piece at a time from a product.

* Comment: since the sales_amount represents the amount paid within an order for a product(independently from the quantity), it would be more reasonable to look at the whole amount paid within an order.
* Comment: from the summary, we see, that the minimum of the quantity and the sales_amount is negative. We won't examing this case or other possibilities of false observations (for example: duplicates, 0 quantity or amount), since it has been already clarified.


### Order frequency
```{r}
purchases %>%
  dplyr::select(order_id, purchase_date) %>%
  group_by(purchase_date) %>%
  summarise(N = length(unique(order_id))) %>%
  ggplot() +
  geom_line(aes(purchase_date, N), stat = "identity")
```


### Quantities of products within an order (with at least 2 hits)
```{r}
purchases %>%
  dplyr::select(quantity) %>%
  group_by(quantity) %>%
  summarise(N = n()) %>%
  filter(N > 1) %>%
  ggplot() +
  geom_bar(aes(quantity, N), stat = "identity")
```


### Sales amount of orders categorized
```{r}
purchases %>%
  dplyr::select(order_id, sales_amount) %>%
  group_by(order_id) %>%
  mutate(
    order_amount = sum(sales_amount),
    order_amount_cat = order_amount %>%
      cut(
        c(0, 50, 100, 200, 500, 1000),
        include.lowest = T
      )
  ) %>%
  group_by(order_amount_cat) %>%
  summarise(N = n()) %>%
  ggplot() +
  geom_bar(aes(order_amount_cat,  N), stat = "identity")
```


### (Weighted) Differences in sales amount of products/piece 
```{r}
purchases %>%
  dplyr::select(product_id, quantity, sales_amount) %>%
  filter(quantity > 0 & sales_amount > 0) %>%
  mutate(product_price = sales_amount / quantity) %>%
  group_by(product_id) %>%
  summarise(
    quantity = sum(quantity),
    avg_price = mean(product_price),
    diff_price = (max(product_price) - min(product_price)) / avg_price
  ) %>%
  ggplot(aes(quantity, diff_price)) +
  geom_jitter() +
  geom_smooth()
```



## Overview of the frequency and amount of purchases (without taking in account the customer or product)

Since we have seen, that there is a trend in the frequency of purchases, it is likely, that there will be also a trend in the amount 
Since we don't consider the customer or product in this overview, we won't need the related columns.
```{r}
sales_raw <- purchases %>%
  dplyr::select(order_id, purchase_date, sales_amount) %>%
  group_by(order_id, purchase_date) %>%
  summarise(sales_amount = sum(sales_amount)) %>%
  ungroup()
```


Calculate the amount of purchases /year day, /weekday, /month, /year
```{r}
sales <- sales_raw %>%
  dplyr::select(sales_amount, purchase_date) %>%
  group_by(purchase_date) %>%
  summarise(total_sales = sum(sales_amount)) %>%
  ungroup() %>%
  mutate(
    month = months(purchase_date),
    month = factor(
      month,
      levels = c(
        "január",
        "február",
        "március",
        "április",
        "május",
        "június",
        "július",
        "augusztus",
        "szeptember",
        "október",
        "november",
        "december"
      )
    ),
    weekday = weekdays(purchase_date),
    weekday = factor(
      weekday,
      levels = c(
        "hétfő",
        "kedd",
        "szerda",
        "csütörtök",
        "péntek",
        "szombat",
        "vasárnap"
      )
    ),
    week = isoweek(purchase_date),
    year = factor(year(purchase_date)),
    day_num = yday(purchase_date)
  )
```


### Daily amount of purchases compared by year
We can see an overall decrease in 2013, with having the same trend within a year.
```{r}
ggplot(data = sales, aes(
  x = day_num,
  y = total_sales,
  group = year,
  color = year
)) +
  geom_line() +
  geom_jitter() +
  geom_smooth(n = 10, span = 0.5)
```


### Sales amount per days of week, weeks of year and months of year

* Days of week: We can see, that customers spent more in the middle of the work week and spent the less on Saturdays
* Weeks of year/months of year: customers spent the most before christmas(week 47-50 / end of november, beginning of december)
```{r sales_trend, fig.height = 11}
sales_weekdays <- sales %>%
  group_by(year, weekday) %>%
  summarise(avg_sales = weighted.mean(total_sales)) %>%
  rename(grouping = weekday)

sales_weeks <- sales %>%
  group_by(year, week) %>%
  summarise(avg_sales = weighted.mean(total_sales)) %>%
  rename(grouping = week)

sales_months <- sales %>%
  group_by(year, month) %>%
  summarise(avg_sales = weighted.mean(total_sales)) %>%
  rename(grouping = month)

sales_weekdays_plot <-
  ggplot(sales_weekdays, aes(x = grouping, y = avg_sales)) +
  geom_bar(stat = "identity", width = 0.5) + xlab("")  + facet_grid(. ~ year)
sales_weeks_plot <- sales_weekdays_plot %+% sales_weeks +
  scale_x_continuous(breaks = seq(min(sales_weeks$grouping), max(sales_weeks$grouping), by = 4))
sales_months_plot <- sales_weekdays_plot %+% sales_months +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 10))

grid.arrange(sales_weekdays_plot,
             sales_weeks_plot,
             sales_months_plot,
             ncol = 1)
```
