---
title: "Ensemble Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 11, fig.height = 11)
```

As a last approach, we will try to fit an ensemble:
In this ensemble, we will use both linear and nonlinear models along with kknn too, hoping to fit the non-linear relationships, what we have seen from earlier and push the model to separate customers, who didn't spent next year from the others
We will also use some preprocessing, like scaling and removing near zero variance predictors and clean the dataset from too outstanding values before training

```{r message = FALSE, warning = FALSE}
library(MASS)
library(car)
library(kknn)
library(mfp)
library(splines)
library(pscl)

source("R/utils.R")
source("R/RFM.R")
source("R/featureScaling.R")
source("R/modelFitting.R")

# Load customer rfm values
customer_rfm_2012 <- loadRFM(2012)
customer_rfm_2013 <- loadRFM(2013)

# Customer data, of spendings in 2012
customer_total_sales_2013 <- customer_rfm_2013 %>% 
  dplyr::select(contact_id, total_sales) %>%
  rename(nexty_total_sales = total_sales)

customers_2012 <- left_join(customer_rfm_2012, customer_total_sales_2013, by = "contact_id") %>% 
  mutate_each(funs(replace(., which(is.na(.)), 0)))
summary(customers_2012)
```

## Preprocess data
###Remove near zero covariates and scale features

```{r}
print(pp <- preProcess(customers_2012[, -c(1, 7)], 
                       method = c("nzv", "scale"),
                       outcome = customers_2012[, 7]))
```

We can see, that active days was discarded and from the plot, that features are now scaled to a smaller range

```{r}
pp$method

pp_customers_2012 <- data.frame(contact_id = customers_2012[, 1],
                                 predict(pp, customers_2012[, -c(1,7)]),
                                 nexty_total_sales = customers_2012[, 7])

plotRfms(pp_customers_2012, c(100, 100, 100, 100, 100, 100), "rfmbn", "Customers preprocessed")
```

### Discarde points with too high leverage, influence 
(in a linear model)

```{r}
set.seed(32)
glm_model <- glm(nexty_total_sales ~ order_num + product_num + total_sales + days_since_last_purchase,
                    data = pp_customers_2012)

lev <- hat(model.matrix(glm_model))
plot(lev)
high_leverage_points <- as.numeric(row.names(pp_customers_2012[lev > 0.15,]))
points(high_leverage_points, lev[high_leverage_points], col = 'red')

cook <- cooks.distance(glm_model)
plot(cook)
points(high_leverage_points, cook[high_leverage_points], col = 'red')
high_influence_points <- as.numeric(row.names(pp_customers_2012[cook > 150,]))

outliners <- union(high_influence_points, high_leverage_points)
pp_customers_2012[outliners,]

pp_customers_2012 <- pp_customers_2012[-outliners,]
```

## Train
For the ensemble we will use the following 4 models:

* Gaussian linear model
* Multiple Fractional Polynomial model
* Polynomial Splines
* Weighted k-Nearest Neighbor Regression model (for 2 clusters)

```{r}
set.seed(3456)
trainIndex <- createDataPartition(pp_customers_2012$nexty_total_sales, p = .8, 
                                  list = FALSE, 
                                  times = 1)

customer_purchase_train <- pp_customers_2012[ trainIndex,]
customer_purchase_test  <- pp_customers_2012[-trainIndex,]
```

#### Quasi-Poisson linear model

```{r}
set.seed(32)
summary(glm_model <- glm(nexty_total_sales ~ order_num + product_num + total_sales + days_since_last_purchase,
                         data = customer_purchase_train))
```

#### Multiple Fractional Polynomial model

```{r}
set.seed(32)
summary(mfp_model <- mfp(nexty_total_sales ~ fp(order_num) + fp(product_num) + fp(total_sales) + fp(days_since_last_purchase), 
                         data = customer_purchase_train))
```

#### Polynomial Splines

```{r}
set.seed(32)
summary(spline_model <- lm(nexty_total_sales ~ bs(order_num) + bs(product_num) + bs(total_sales) + bs(days_since_last_purchase), 
                         data = customer_purchase_train))
```

#### Weighted k-Nearest Neighbor Regression model

```{r}
set.seed(32)
summary(kknn_model <- train.kknn(nexty_total_sales ~ order_num + product_num + total_sales + days_since_last_purchase, 
                                 data = customer_purchase_train,
                                 ks = c(2)))
```

### Predict test data
```{r}
testData_glm <- cbind(customer_purchase_test, fit = predict(glm_model, customer_purchase_test, type = "response"))
testData_mfp <- cbind(customer_purchase_test, fit = predict(mfp_model, customer_purchase_test, type = "response"))
testData_spline <- cbind(customer_purchase_test, fit = predict(spline_model, customer_purchase_test))
testData_kknn <- cbind(customer_purchase_test, fit = predict(kknn_model, customer_purchase_test))
```

#### Get weights for ensemble with minimizing the RMSE
The ensebmle optimizes for RMSE with a weighted mean calculation of the training results of each model

```{r}
set.seed(32)
ens_weights <- optimiseEnsemble(fitted1 = glm_model$fitted.values, 
                                fitted2 = mfp_model$fitted.values,
                                fitted3 = spline_model$fitted.values,
                                fitted4 = unlist(kknn_model$fitted.values),
                                actual = glm_model$y)$par
ens_weights
```

### Calculating differences in predicted and actual data
We can see, that all models did pretty good (compared to earlier models) both regarding the RMSE and the actual difference

```{r}
glm_test_RMSE <- calculateRMSEFromFit(testData_glm)
glm_test_diff <- calculateDiffFromModel(testData_glm)
glm_test_diff_perc <- calculateDiffFromModel(testData_glm) * 100 / sum(testData_glm$nexty_total_sales)
print(paste("RMSE:", glm_test_RMSE, ", diff:", glm_test_diff, ",diff(%):", glm_test_diff_perc))

mfp_test_RMSE <- calculateRMSEFromFit(testData_mfp)
mfp_test_diff <- calculateDiffFromModel(testData_mfp)
mfp_test_diff_perc <- calculateDiffFromModel(testData_mfp) * 100 / sum(testData_mfp$nexty_total_sales)
print(paste("RMSE:", mfp_test_RMSE, ", diff:", mfp_test_diff, ",diff(%):", mfp_test_diff_perc))

spline_test_RMSE <- calculateRMSEFromFit(testData_spline)
spline_test_diff <- calculateDiffFromModel(testData_spline)
spline_test_diff_perc <- calculateDiffFromModel(testData_spline) * 100 / sum(testData_spline$nexty_total_sales)
print(paste("RMSE:", spline_test_RMSE, ", diff:", spline_test_diff, ",diff(%):", spline_test_diff_perc))

kknn_test_RMSE <- calculateRMSEFromFit(testData_kknn)
kknn_test_diff <- calculateDiffFromModel(testData_kknn)
kknn_test_diff_perc <- calculateDiffFromModel(testData_kknn) * 100 / sum(testData_kknn$nexty_total_sales)
print(paste("RMSE:", kknn_test_RMSE, ", diff:", kknn_test_diff, ",diff(%):", kknn_test_diff_perc))
```

From the results we can see, that the ensemble did also pretty good both regarding the RMSE and especially in the accuracy in predicting the next years revenue

```{r}
testData_ensemble <- data.frame(testData_glm[, c(1:6)],
                                fit = ensembleFun(testData_glm$fit, testData_mfp$fit, testData_spline$fit, testData_kknn$fit, ens_weights))
ens_test_RMSE <- calculateRMSEFromFit(testData_ensemble)
ens_test_diff <- calculateDiffFromModel(testData_ensemble)
ens_test_diff_perc <- calculateDiffFromModel(testData_ensemble) * 100 / sum(testData_ensemble$nexty_total_sales)
print(paste("RMSE:", ens_test_RMSE, ", diff:", ens_test_diff, ",diff(%):", ens_test_diff_perc))
```

Next, we will train for the whole dataset and test the ensemble for some customer groups

### Train on whole dataset

```{r}
set.seed(32)
final_model_glm <-glm(nexty_total_sales ~ order_num + product_num + total_sales + days_since_last_purchase,
                      data = pp_customers_2012)
set.seed(32)
final_model_mfp <- mfp(nexty_total_sales ~ fp(order_num) + fp(product_num) + fp(total_sales) + fp(days_since_last_purchase), 
                        data = pp_customers_2012)
set.seed(32)
final_model_spline <- lm(nexty_total_sales ~ bs(order_num) + bs(product_num) + bs(total_sales) + bs(days_since_last_purchase), 
                                data = pp_customers_2012)
set.seed(32)
final_model_kknn <- train.kknn(nexty_total_sales ~ order_num + product_num + total_sales + days_since_last_purchase, 
                                       data = pp_customers_2012,
                                       ks = c(2))
```

### Test customer groups
We will test on the same two customer groups as before: customers who purchased in december and "onetime" customers (order_num == 1)

#### Dec customers
Filter december customers, predict next years revenue and optimise ensemble weights

```{r}
pp_dec_customers <- pp_customers_2012 %>%
  filter(days_since_last_purchase < 31)

decData_fitted_glm <- cbind(pp_dec_customers, fit = predict(final_model_glm, pp_dec_customers, type = "response"))
decData_fitted_mfp <- cbind(pp_dec_customers, fit = predict(final_model_mfp, pp_dec_customers, type = "response"))
decData_fitted_spline <- cbind(pp_dec_customers, fit = predict(final_model_spline, pp_dec_customers))
decData_fitted_kknn <- cbind(pp_dec_customers, fit = predict(final_model_kknn, pp_dec_customers))

ens_weights <- optimiseEnsemble(fitted1 = final_model_glm$fitted.values, 
                                fitted2 = final_model_mfp$fitted.values,
                                fitted3 = final_model_spline$fitted.values,
                                fitted4 = unlist(final_model_kknn$fitted.values),
                                actual = final_model_glm$y)$par
ens_weights
```

Calculate ensemble result and calculate error, accuracy

```{r}
decData_fitted_ens <- data.frame(decData_fitted_glm[, c(1:6)],
                                fit = ensembleFun(decData_fitted_glm$fit, decData_fitted_mfp$fit, decData_fitted_spline$fit, decData_fitted_kknn$fit, 
                                                  ens_weights))

# Sum of differences, RMSE and % in dec customers
ens_dec_RMSE <- calculateRMSEFromFit(decData_fitted_ens)
ens_dec_diff <- calculateDiffFromModel(decData_fitted_ens)
ens_dec_diff_perc <- calculateDiffFromModel(decData_fitted_ens) * 100 / sum(decData_fitted_ens$nexty_total_sales)
print(paste("RMSE:", ens_dec_RMSE, ", diff:", ens_dec_diff, ",diff(%):", ens_dec_diff_perc))

```

We can see, that the ensemble model had even better results on this customer group with an almost 100% accuracy :)

### "Onetime" customers
Filter onetime customers and predict next years revenue

```{r}
order_min_scaled <- min(pp_customers_2012$order_num)
onetime_customers <- pp_customers_2012 %>%
  filter(order_num == order_min_scaled)


onetimeData_fitted_glm <- cbind(onetime_customers, fit = predict(final_model_glm, onetime_customers, type = "response"))
onetimeData_fitted_mfp <- cbind(onetime_customers, fit = predict(final_model_mfp, onetime_customers, type = "response"))
onetimeData_fitted_spline <- cbind(onetime_customers, fit = predict(final_model_spline, onetime_customers))
onetimeData_fitted_kknn <- cbind(onetime_customers, fit = predict(final_model_kknn, onetime_customers))
```

Calculate ensemble result and calculate error, accuracy

```{r}
onetimeData_fitted_ens <- data.frame(onetimeData_fitted_glm[, c(1:6)],
                                 fit = ensembleFun(onetimeData_fitted_glm$fit, onetimeData_fitted_mfp$fit, onetimeData_fitted_spline$fit, onetimeData_fitted_kknn$fit,
                                                   ens_weights))

# Sum of differences, RMSE and % in onetime customers
ens_onetime_RMSE <- calculateRMSEFromFit(onetimeData_fitted_ens)
ens_onetime_diff <- calculateDiffFromModel(onetimeData_fitted_ens)
ens_onetime_diff_perc <- calculateDiffFromModel(onetimeData_fitted_ens) * 100 / sum(onetimeData_fitted_ens$nexty_total_sales)
print(paste("RMSE:", ens_onetime_RMSE, ", diff:", ens_onetime_diff, ",diff(%):", ens_onetime_diff_perc))
```

**So far, this is the best model, so we will choose it for prediction and create an interface for accessing it and predict new data with it**

## Save model and related data
Save preprocessed customer data for both 2012 and 2013

```{r}
pp_customers_2012 <- data.frame(contact_id = customers_2012[, 1],
                                predict(pp, customers_2012[, -c(1,7)]),
                                nexty_total_sales = customers_2012[, 7])
write.csv(pp_customers_2012, "data/RFM_pp_2012.csv", row.names = F, quote = F)

pp_customers_2013 <- data.frame(contact_id = customer_rfm_2013[, 1],
                                predict(pp, customer_rfm_2013[, -1]))
write.csv(pp_customers_2013, "data/RFM_pp_2013.csv", row.names = F, quote = F)
```

Save models and ensemble weights

```{r}
saveRDS(final_model_glm, file = "data/model/final_glm.rds")
saveRDS(final_model_mfp, file = "data/model/final_mfp.rds")
saveRDS(final_model_spline, file = "data/model/final_spline.rds")
saveRDS(final_model_kknn, file = "data/model/final_kknn.rds")
saveRDS(ens_weights, file = "data/model/ens_weights.rds")
```

## Save results

```{r}
glm_results <- createModelSummary(description = "Gaussian linear regression, scaled features, formula = nexty_total_sales ~ order_num + product_num + total_sales + days_since_last_purchase",
                                  test_RMSE = glm_test_RMSE,
                                  test_diff = glm_test_diff,
                                  test_diff_percent = glm_test_diff_perc)
saveRDS(glm_results, file = "data/results/glm.rds")

mfp_results <- createModelSummary(description = "Multiple Fractional Polynomial model, scaled features, formula = nexty_total_sales ~ order_num + product_num + total_sales + days_since_last_purchase",
                                  test_RMSE = mfp_test_RMSE,
                                  test_diff = mfp_test_diff,
                                  test_diff_percent = mfp_test_diff_perc)
saveRDS(mfp_results, file = "data/results/mfp.rds")

spline_results <- createModelSummary(description = "Polynomial Splines, scaled features, formula = nexty_total_sales ~ order_num + product_num + total_sales + days_since_last_purchase",
                                    test_RMSE = spline_test_RMSE,
                                    test_diff = spline_test_diff,
                                    test_diff_percent = spline_test_diff_perc)
saveRDS(spline_results, file = "data/results/spline.rds")

kknn_results <- createModelSummary(description = "Polynomial Splines, scaled features, formula = nexty_total_sales ~ order_num + product_num + total_sales + days_since_last_purchase",
                                   test_RMSE = kknn_test_RMSE,
                                   test_diff = kknn_test_diff,
                                   test_diff_percent = kknn_test_diff_perc)
saveRDS(kknn_results, file = "data/results/kknn.rds")

ens_results <- createModelSummary(description = "Ensemble of (gml, mfp, splines), formula = nexty_total_sales ~ order_num + product_num + total_sales + days_since_last_purchase",
                                   test_RMSE = ens_test_RMSE,
                                   test_diff = ens_test_diff,
                                   test_diff_percent = ens_test_diff_perc,
                                  dec_test_RMSE = ens_dec_RMSE,
                                  dec_test_diff = ens_dec_diff,
                                  dec_test_diff_percent = ens_dec_diff_perc,
                                  onetime_test_RMSE = ens_onetime_RMSE,
                                  onetime_test_diff = ens_onetime_diff,
                                  onetime_test_diff_percent = ens_onetime_diff_perc)
saveRDS(ens_results, file = "data/results/ens.rds")
```
