---
title: "Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will have a quick overview of the results and an example on how to use the fitted model.

## Results of models

```{r message = FALSE, warning = FALSE}
source("R/utils.R")

model_results <- loadResults()
printModelResults(model_results)

```

## Revenue predictor example
The functions for testing on 2012 -> 2013 customer spending and predicting 2013 -> 2014 customer spending can be found in the root folder in the "RevenuePredictor.R" file

```{r message = FALSE, warning = FALSE}
library(lubridate)
source("RevenuePredictor.R")

# Get customers, who purchased in 2012 and 2013 november
group2012 <- purchases %>%
  filter(year(purchase_date) == 2012 & month(purchase_date) == 11) %>%
  select(contact_id) %>%
  unique
group_2013 <- purchases %>%
  filter(year(purchase_date) == 2013 & month(purchase_date) == 11) %>%
  select(contact_id) %>%
  unique

# Test
group1_test_results <- testRevenue(group2012$contact_id)
summary(group1_test_results)


# Predict
group1_pred_results <- predictRevenue(group_2013$contact_id)
summary(group1_pred_results)
```

We can see, that the prediction contained some pretty big differences, because in this dataset, the outliner customers are not filtered and the ensemble model is not prepared for them... 

### Some final toughts

* With filtering the outstanding points, the model became pretty accurate, but also "unstable", if a datapoint is not in the normally fitting range, it can make huge errors.. even predict negative revenue.
* On the other hand, if we did not filter the outstanding points, the model couldn't fit so good for the non-outstanding data.
* For resolving the negative prediction, I also tried adding a quasi-poisson fitting (it's not included in the code), but it turned out to be unaccurate, so I didn't included it in the model.
* If this revenue prediction problem has a solution, I think, it includes separating these "high-end" customers from the others and also those customers, who were spending a lot this year, but didn't purchased anything next year


